---
title: "HW2_JeromeWong"
author: "Jerome Wong"
date: "11/10/2021"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    number_sections: no
---

```{r setup, include=FALSE}
library(igraph)
knitr::opts_chunk$set(echo = TRUE)
```


```{r, fig.align='center'}
# Define our graph
dg <- graph.formula(1++2, 1++3, 2++3, 2++4)
# Plot; permute changes the orientation shown in the plot
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```

### Spillover
### Qn1

```{r, fig.align='center'}
# assign treatment labels T (treatment) or C (control)
#  upper left, upper right, lower left, lower right
V(dg)$label <- c("T", "T", "C", "C")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 0 no exposure
(b) 0 direct exposure
(c) 2 indirect exposure
(d) 2 full exposure


\newpage
```{r}
V(dg)$label <- c("T", "C", "T", "C")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 1 no exposure
(b) 0 direct exposure
(c) 1 indirect exposure
(d) 2 full exposure


\newpage
```{r}
V(dg)$label <- c("T", "C", "C", "T")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 0 no exposure
(b) 2 direct exposure
(c) 2 indirect exposure
(d) 0 full exposure


\newpage
```{r}
V(dg)$label <- c("C", "T", "T", "C")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 0 no exposure
(b) 0 direct exposure
(c) 2 indirect exposure
(d) 2 full exposure


\newpage
```{r}
V(dg)$label <- c("C", "T", "C", "T")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 0 no exposure
(b) 0 direct exposure
(c) 2 indirect exposure
(d) 2 full exposure


\newpage
```{r}
V(dg)$label <- c("C", "C", "T", "T")
plot(permute(dg, c(3,4,1,2)), layout=layout.grid(dg))
```
(a) 0 no exposure
(b) 2 direct exposure
(c) 2 indirect exposure
(d) 0 full exposure


\newpage
### Qn2

This would violate the second assumption of the the Horvitz-Thompson estimators. The observed in-degree of the network would be incorrect and since the in-degree is a function of the probabilities in the HT estimators, the probabilities obtained would be wrong. The inaccuracies in measuring the network introduces bias into the HT estimators.
